Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-05-27 Manuscript/01-Intro.Rmd

#### Spike-and-slab prior
The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior. It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
$$
 \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
$$

@@

In this dissertation, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model. Specifically, we will propose a new Bayesian spike-and-slab prior for the generalized additive models and Cox spline models. The contribution of the proposed prior comes in two-fold: 1) it supports simultaneous variable selection and function selection; 2) it provides robust estimation and prediction for the coefficients and outcomes. As previously described, simultaneous variable selection and its functional form selection is not a trivial problem. Nevertheless, the Bayesian spike-and-slab prior is a perfect solution to the problem, as the latent binary indicator can be used to for variable selection and the spike and slab densities are used to adaptive adjust the smoothness of the spline function. Moreover, compared to currently available Bayesian spline models which mostly conduct a 'all-in-all-out' strategy, i.e. either including a non-linear form of a variable or excluding the variable completely, we offer a more flexible solution, rending one of three possibilities: no effect of a predictor, only linear effect of a predictor, or non-linear effect. This simplifies the process of determining what effect a predictor have on the outcome, without involving hypothesis testing where the degree of freedom for non-linear effect is complicated to estimate due to the shrinkage. Moreover, out model inherent the the robust estimation from Bayesian models.

>>>>

For high-dimensional data analysis, a common Bayesian modeling approach is to impose a particular family of priors, the spike-and-slab (SS) priors on the coeffcient in a regression modelss. The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior. It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
$$
 \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
$$

@@

In this article, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model. Specifically, we will propose a new Bayesian spike-and-slab prior for the generalized additive models and Cox spline models. The contribution of the proposed prior comes in two-fold: 1) it supports simultaneous variable selection and function selection; 2) it provides robust estimation and prediction for the coefficients and outcomes. As previously described, simultaneous variable selection and its functional form selection is not a trivial problem. Nevertheless, the Bayesian spike-and-slab prior is a perfect solution to the problem, as the latent binary indicator can be used to for variable selection and the spike and slab densities are used to adaptive adjust the smoothness of the spline function. Moreover, compared to currently available Bayesian spline models which mostly conduct a 'all-in-all-out' strategy, i.e. either including a non-linear form of a variable or excluding the variable completely, we offer a more flexible solution, rending one of three possibilities: no effect of a predictor, only linear effect of a predictor, or non-linear effect. This simplifies the process of determining what effect a predictor have on the outcome, without involving hypothesis testing where the degree of freedom for non-linear effect is complicated to estimate due to the shrinkage. Moreover, out model inherent the the robust estimation from Bayesian models.
